# The climate of equality, diveristy and inclusion

 
## Perceptions of gender equality

One of the consistent findings of the ASSET Survey 2016 is that men reported a more positive perception of Equality, Diversity and Inclustion aspects than women. The GEAM captures the perception of gender equality issues through several items. `OCPER001` inquires about any differences in organizational committment to gender equality or differences in the equal treatment and participation of women and men. `OCPER003` on the other hand examines differences in the allocation of several type of resources and responsibilities. 


+------------------+--------------------------------------------------------------------------------------------------+
| GEAM Code        | Item                                                                                             |
+==================+==================================================================================================+
| OCPER.SQ001.     | In general, men and women are equally represented (in terms of numbers) in my `VarOrgType.shown` | 
+------------------+--------------------------------------------------------------------------------------------------+
| OCPER.SQ002.     | In general, men and women are treated equally in my`VarOrgType.shown`                            |
+------------------+--------------------------------------------------------------------------------------------------+
| OCPER.SQ003.     | My  `VarOrgType.shown` is committed to promoting gender equality                                 |
+------------------+--------------------------------------------------------------------------------------------------+
| OCPER.SQ004.     | Myself and colleagues know who to go to if we have concerns related to gender equality           | 
+------------------+--------------------------------------------------------------------------------------------------+
| OCPER.SQ005.     | My  `VarOrgType.shown` is responsive to concerns about gender equality                           | 
+------------------+--------------------------------------------------------------------------------------------------+

Responses to `OCPER001` are coded on a Likerst scale from "1-Strongly disasgree" to "5-Strongly Agree". 

```{r ocper-ge, fig.cap="Perceptions of gender equality", out.width="100%", fig.align='center'}

cpal <- RColorBrewer::brewer.pal(5, "Set1")

df.geam %>% 
  select(starts_with("OCPER001.SQ")) %>% 
  plot_stackfrq(geom.colors =cpal,
                geom.size = .4,
                show.total = F, 
                digits = 0, 
                show.axis.prc = T) +
  theme_light() + 
  theme(legend.position = "bottom", 
        legend.direction = "horizontal",
        legend.justification = c(1,0),
        plot.margin = unit(c(.2,2,.2,.2), "cm"), #t,r,b,l
        axis.text.x=element_blank(),
        axis.text = element_text(size = 10))


```


### Preparation and pre-processing

We can convert the answer options from a categorical factor variable to a continous numeric variable, where agreement ranges from 1 to 
5.

```{r}

df.geam %<>%
    mutate(OCPER001.RE.SQ001. = as.numeric(OCPER001.SQ001.), 
           OCPER001.RE.SQ002. = as.numeric(OCPER001.SQ002.), 
           OCPER001.RE.SQ003. = as.numeric(OCPER001.SQ003.),
           OCPER001.RE.SQ004. = as.numeric(OCPER001.SQ004.), 
           OCPER001.RE.SQ005. = as.numeric(OCPER001.SQ005.))
```


We'll create a subset including a binary `gender` variable and `ethnic minority status`, eliminating al `NA` responses. 

```{r}
df.sub <- df.geam %>% 
  filter(!is.na(SDEM004.bin)) %>% 
  select(starts_with("OCPER001."), Gender=SDEM004.bin, Minority=SDEM002)
```




### Differing perceptions by gender? 

First, let's look at the mean values for each of the response items by gender. Higher average values indicate a stronger agreement with the specified sub-question. 

```{r eval=F, echo=F}
df.sub %>%
    pivot_longer(cols=starts_with("OCPER001.RE."), names_to="Subquestion", values_to="Perception") %>% 
    ggplot(aes(x=Subquestion, y=Perception, fill=Gender)) + 
    geom_boxplot() + 
#    stat_summary(fun.y=mean, shape=5, size=1, color="red", fill="red") +
    scale_fill_manual(values=cpal) +
    labs(x="") +
    theme_light() + 
    theme(legend.position = "right", 
        #axis.text.x=element_blank(),
        axis.text = element_text(size = 10))

```

```{r}

tbl1 <- df.sub %>% 
    group_by(Gender) %>% 
    summarize(N = n(),
              "Equal representation"         = mean(OCPER001.RE.SQ001., na.rm=T), 
              "Equal treatment"              = mean(OCPER001.RE.SQ002., na.rm=T),
              "Org. is committed"            = mean(OCPER001.RE.SQ003., na.rm=T), 
              "Know equality contact person" = mean(OCPER001.RE.SQ004., na.rm=T), 
              "Org. is responsive"           = mean(OCPER001.RE.SQ005., na.rm=T)) %>% 
    pivot_longer(names_to = "Perception", values_to = "Mean", cols=c(3:7)) %>% 
    arrange(Perception)

```

```{r echo=F}
kbl(tbl1, 
    booktabs = T,
    digits=3, 
    caption="Mean perceptions of gender equality") %>%
    kable_styling(full_width=T, latex_options =c("hold_position"))
```

As the data indicates, men on average agree stronger the both genders are equally represented in the organisation, while women do not share this opinion. Women preceive that they are under-represented in the organization. The same holds for the perception of equal treatment, where men thing that both genders are treated equally while women tend to disagree more. There is also a difference on `OCPER001.SQ004` ("Myself and colleagues know who to go to if we have concerns related to gender equality") with women indicating that they agree more strongly than men. 


In order to decide if there are indeed, statistically significant gender differences for each of the available sub-question items of  `OCPER001` we carry out a [ANOVA](#anova): 

```{r eval=F}
sq01.gender <- aov(OCPER001.RE.SQ001. ~ Gender, data=df.sub)
sq02.gender <- aov(OCPER001.RE.SQ002. ~ Gender, data=df.sub)

```

... etc. 

```{r echo=F}

# create all anovas
sq01.gender <- aov(OCPER001.RE.SQ001. ~ Gender, data=df.sub)
sq02.gender <- aov(OCPER001.RE.SQ002. ~ Gender, data=df.sub)
sq03.gender <- aov(OCPER001.RE.SQ003. ~ Gender, data=df.sub)
sq04.gender <- aov(OCPER001.RE.SQ004. ~ Gender, data=df.sub)
sq05.gender <- aov(OCPER001.RE.SQ005. ~ Gender, data=df.sub)

# combine individual results into one table
ocper001.gender <- broom::tidy(sq01.gender) %>% 
    add_row(broom::tidy(sq02.gender)) %>% 
    add_row(broom::tidy(sq03.gender)) %>% 
    add_row(broom::tidy(sq04.gender)) %>% 
    add_row(broom::tidy(sq05.gender)) 
```

The overall output looks like follows: 

```{r echo=F}
kbl(ocper001.gender, 
    booktabs = T, 
    digits=3, 
    caption="Anova: perception of equality by gender") %>%
    group_rows(group_label="Equal representation", start_row=1, end_row=2) %>%
    group_rows(group_label="Equal treatment", start_row=3, end_row=4) %>%
    group_rows(group_label="Org. is committed", start_row=5, end_row=6) %>%
    group_rows(group_label="Know equality contact person", start_row=7, end_row=8) %>%
    group_rows(group_label="Org. is responsive", start_row=9, end_row=10) %>%
    kable_styling(full_width=T, latex_options =c("hold_position", "scale_down"))
```



\newpage 



### Differing perceptions by gender and minority status

For the calculation of the minority status, we only consider response options "Majory" and "Minority" while removing "Unsure" and "Prefer not to say". 

```{r echo=F}

df.sub <- df.geam %>% 
  filter(!is.na(SDEM004.bin) & (SDEM002 == "Majority" | SDEM002 == "Minority")) %>% 
  select(starts_with("OCPER001."), Gender=SDEM004.bin, Minority=SDEM002)

```

Let's examine the mean values of each `OCPER001` sub-question according to gender and minority status: 

```{r}

tbl2 <- df.sub %>% 
    group_by(Gender, Minority) %>% 
    summarize(N = n(),
              "Equal representation"         = mean(OCPER001.RE.SQ001., na.rm=T), 
              "Equal treatment"              = mean(OCPER001.RE.SQ002., na.rm=T),
              "Org. is committed"            = mean(OCPER001.RE.SQ003., na.rm=T), 
              "Know equality contact person" = mean(OCPER001.RE.SQ004., na.rm=T), 
              "Org. is responsive"           = mean(OCPER001.RE.SQ005., na.rm=T)) %>% 
    pivot_longer(names_to = "Perception", values_to = "Mean", cols=c(4:8)) %>% 
    arrange(Perception, Minority)
    
```

```{r echo=F}
kbl(tbl2, 
    booktabs = T,
    digits=3, 
    caption="Mean perceptions of equality by minority status and gender") %>%
    kable_styling(full_width=T, latex_options =c("hold_position", "scale_down"))
```





```{r echo=F}

# create all anovas
sq01.gemin  <- aov(OCPER001.RE.SQ001. ~ Gender + Minority + Gender * Minority, data=df.sub)
sq02.gemin  <- aov(OCPER001.RE.SQ002. ~ Gender + Minority + Gender * Minority, data=df.sub)
sq03.gemin  <- aov(OCPER001.RE.SQ003. ~ Gender + Minority + Gender * Minority, data=df.sub)
sq04.gemin  <- aov(OCPER001.RE.SQ004. ~ Gender + Minority + Gender * Minority, data=df.sub)
sq05.gemin  <- aov(OCPER001.RE.SQ005. ~ Gender + Minority + Gender * Minority, data=df.sub)

# combine individual results into one table
ocper001.gemin <- broom::tidy(sq01.gemin) %>% 
    add_row(broom::tidy(sq02.gemin)) %>% 
    add_row(broom::tidy(sq03.gemin)) %>% 
    add_row(broom::tidy(sq04.gemin)) %>% 
    add_row(broom::tidy(sq05.gemin)) 
```



The overall output looks like follows: 

```{r echo=F}
kbl(ocper001.gemin, 
    booktabs = T, 
    digits=3, 
    caption="Anova: perception of gender equality by gender") %>%
    group_rows(group_label="Equal representation", start_row=1, end_row=4) %>%
    group_rows(group_label="Equal treatment", start_row=5, end_row=8) %>%
    group_rows(group_label="Org. is committed", start_row=9, end_row=12) %>%
    group_rows(group_label="Know equality contact person", start_row=13, end_row=16) %>%
    group_rows(group_label="Org. is responsive", start_row=17, end_row=20) %>%
    kable_styling(full_width=T, latex_options =c("hold_position"))
```



\newpage

